### 2004-06-28 doc/Simulation_and_Control_using_Supremica/introduction.tex
*** empty log message ***

\chapter{Introduction}
\label{introduction}

Define the problem.
Define other approaches, state of the art.
Introduce Function Blocks
Introduce Automation Objects
(Configuration, Simulation, Control, HMI)