### 2004-07-01 doc/Simulation_and_Control_using_Supremica/automationobjects.tex
*** empty log message ***

\chapter{Automation Objects}
\label{automationobjects}

%Introduce Automation Objects

%% Why automation objects? Background.
Using only function blocks efficient hierarchical control
system design is very time consumig. To facilitate easier
and more rapid design process a new kind of design object
needs to be introduced. Before that the purpose of that kind
of object has to be defined.

%% Definition. Multi view objects where the view is the
%% interface exposed for different purposes.

% Define the different views (Model-View-Controller) if relevant.
%% Configuration
%% Human-Machine Interface
%% Control
%% Simulation

% Introduce the example