### 2004-07-16 doc/Simulation_and_Control_using_Supremica/functionblocks.tex

\chapter{IEC~61499 Standard}
\label{functionblocks}

\textbf{This chapter,with minor revisions, is copied from
  chapter 2 in \cite{j:des:2002}.}

New technologies and standards are emerging that are set to
have a dramatic effect on the design and implementation of
industrial control systems. A new standard
IEC~61131-3~\cite{iec:1131:1993} was introduced in 1995 to
standardize the representation of control designs so as to
enable cross-vendor integration. The standard offered
significant advantages over earlier vendor proprietary
control design representations~\cite{l:pro:1995}.
\begin{itemize}
\item It encouraged well-structured procedural programs that
  allowed the development of different functional elements
  into separate procedures or program units.
\item It provided standardized languages and methods of
  program execution so that wide range of technological
  problems can be programmed as vendor-independent software
  elements.
\item It provided support for various languages like
  structured text, function block diagrams, ladder diagrams,
  instruction list and sequential function charts.
\item The standard enforced strong data typing and supported
  user defined data structures.
\end{itemize}

The standard however was based on a scanning mechanism where
the program execution was monitored by a continuous loop.
Although different parts of the program could be scanned at
different frequencies (based on fidelity requirements), it
was tedious to describe such behavior using sequential
function charts. The standard was based on polling mechanism
rather than an event driven system, and hence its execution
order is not clearly defined.

Further the standard was designed to support a software
model and languages for PLCs whose software was typically
running on a single processing resource. The communication
between different resources was supported through global
variables and communication blocks, both of which did not
provide a clear and concise method for defining the
connections between distributed systems.

To overcome these problems, a function block model was
sought that would support distributed control systems in a
scalable, extensible and flexible manner.
IEC~61499-1~\cite{iec:614991:2000} standard has now been
developed for distributed industrial process, measurement
and control systems.

In industrial systems, function blocks are now an
established concept for defining robust, re-usable software
components. A {\it Function Block} instance is defined to a
software function unit comprising an individual, named copy
of a data structure and associated operations specified by a
corresponding function type.

Function blocks are analogous to objects in object-oriented
programming. Control designers can now take standard proven
encapsulated functionality in the form of blocks and link it
together in the quickest and most intuitive way possible.
They reduce the complexity since system integrators can now
build large systems using {\it black-box composition} i.e.
without worrying about how the internals of a block work.

The following sections are adapted largely from the various
discussions on the features of the standard in
~\cite{iec:614991:2000, iec:1131:1993, l:pro:1995,
  l:mod:2001, c:des:2002, c:ope:2002}

\section{Reference Models}
IEC~61499-1 has provided a set of models for describing
distributed systems that have been programmed using function
blocks.

\subsection{System Model} 
At the physical level, a distributed system consists of a
set of co-operating {\it applications}. An application can
exist on a single system such as application C in figure
\ref{f:System_Model} or have functionality distributed over
a number of devices such as applications A and B in Figure
\ref{f:System_Model}.

\begin{figure}
  \begin{center}
    \epsfig{file=images/functionblocks/System_Model.eps,scale=.8} 
    \caption[System Model]
            {System Model{\protect ~\cite{iec:614991:2000}}}
    \label{f:System_Model}
  \end{center}
\end{figure}

\subsection{Device Model} 
A resource provides independent execution and control of
networks of function blocks. It supports zero or more {\it
  resources} and has at least one interface. The interface
could either be a 'process interface' that allows the
resources to exchange data with the input and output points
on the physical devices, or a communication interface that
allows the resources to communicate with external resources
in remote devices. The device model is illustrated in Figure
\ref{f:Device_Model}.

\begin{figure}
  \begin{center}
    \epsfig{file=images/functionblocks/Device_Model.eps,scale=.8} 
    \caption[Device Model]
            {Device Model{\protect ~\cite{iec:614991:2000}}}
    \label{f:Device_Model}
  \end{center}
\end{figure}

\subsection{Resource Model} 
A resource is an independent functional unit, contained in a
device which has independent control of its operation. The
resource can be created, configured, parameterized, started
up, deleted etc. without effecting other resources. Each
resource provides an interface to the communication systems
and tot the device specific process. The resource is
therefore, concerned with the mapping of data and event
flows which pass between function blocks in the local
resource to remove resource function blocks via the device
communication interfaces.

Figure \ref{f:Resource_Model} depicts the main features of
an IEC 61499 resource. Within the resource, it shows a
network of interconnected function blocks linked by data and
event flows. A scheduling function ensures that the events
are executed in the correct serial order. {\it Service
  Interface} function blocks are a special form of function
blocks that provide a link between function blocks and the
interface of the resource.

\begin{figure}
  \begin{center}
    \epsfig{file=images/functionblocks/Resource_Model.eps,scale=.8} 
    \caption[Resource Model]
            {Resource Model{\protect ~\cite{iec:614991:2000}}}
    \label{f:Resource_Model}
  \end{center}
\end{figure}

\subsection{Application Model} 
An application is defined as a network of interconnected
function blocks, linked by data and event flows. An
application can be fragmented and distributed over many
resources. The resources ensure that the events are executed
in the correct serial order based on the causal
relationships specified by the applications.

An application could be understood as an entire set of
function blocks and interconnections to solve a particular
automation control problem. Figure \ref{f:Application_Model}
illustrates an application, where each individual block is
either a function block or a subapplication.

\begin{figure}
  \begin{center}
    \epsfig{file=images/functionblocks/Application_Model.eps, scale=.8}
    \caption[Application Model]
            {Application Model{\protect ~\cite{iec:614991:2000}}} 
    \label{f:Application_Model}
  \end{center}
\end{figure}

\subsection{Function Block Model} 
A function block is a functional unit of software that its
own data structure which can be manipulated by one or more
algorithms. The formal description of the data structure is
provided by the function block type definition.

The main features of function blocks are:
\begin{itemize}
\item Each function block has a type name and an instance
  name.
\item Each block ha a set of event inputs, which can
  receive inputs from other blocks through event
  connections.
\item Each block has one or more outputs, through which it
  can propagate its results to other blocks.
\item There is a set of data inputs that allow values to
  be passed in from other blocks.
\item There is a set of data outputs, through which the
  block can pass its results to other blocks.
\item Each block may have a set of internal variables that
  hold its state between algorithm invocations.
\item The behavior of the block is defined in terms of
  algorithms and state information. This behavior is
  defined in the function block type definition.
\end{itemize}

Figure \ref{f:Function_Block_Char} shows the main
characteristics of a function block. The top part of the
block, called the 'execution control' portion contains in
some cases, a state machine that maps events to the
algorithms that should be triggered as a consequence of the
event. The lower portion of block contains both the
algorithms and the internal data, both of which are
encapsulated within the block.

\begin{figure}
  \begin{center}
    \epsfig{file=images/functionblocks/Function_Block_Char.eps, scale=.8}
    \caption[Function Block Characteristics]
            {Function Block Characteristics{\protect ~\cite{iec:614991:2000}}}
    \label{f:Function_Block_Char}
  \end{center}
\end{figure}

\section{Function Block Types}
A function block type is defined by a type name, formal
definitions for the input events and data and the output
events and data. The type definition also includes the
internal behavior of the block. The notion of a function
block type is analogous to the class definition in
object-oriented programming while function block instances
correspond to the object instances of the classes.

\subsection{Basic Function Block Types} 
The behavior of basic function blocks is defined in terms of
algorithms that are invoked in response to input events. As
algorithms execute, they trigger output events to signal
their state to other blocks. The basic blocks hold their
state in terms of internal variables. Basic function blocks
also can communicate their state using input and output data
variables.

The mapping of events onto algorithms is expressed using a
state transition notation known as execution control chart
and is described in the next section. Figure
\ref{f:Basic_Block} shows an example of a basic function
block.

\begin{figure}
  \begin{center}
    \epsfig{file=images/functionblocks/Basic_Function_Block.eps,scale=.8}
    \caption[Basic Function Block]
            {Basic Function Block{\protect ~\cite{iec:614991:2000}}} 
    \label{f:Basic_Block}
  \end{center}
\end{figure}

\subsection{Composite Function Blocks} 
The behavior of a composite function block is described in
terms of the event and data connections between its
component function blocks. A composite function block can be
used to hide the internal interactions between the component
blocks and presenting a simplified interface for use by the
outside world. A composite function block is illustrated in
Figure \ref{f:Composite_Block}.

\begin{figure}
  \begin{center}
    \epsfig{file=images/functionblocks/Composite_Function_Block.eps, scale=.8}
    \caption[Composite Function Block]
            {Composite Function Block{\protect ~\cite{iec:614991:2000}}} 
        \label{f:Composite_Block}
  \end{center}
\end{figure}

\subsection{Service Interface Function Blocks} 
Service interface function blocks provide an interface
between the function block domain and the external services.
Since a service interface function block is primarily
concerned with data exchange, it is defined using time
sequence diagrams. Examples of service interface function
blocks are communication blocks for publishing and
subscribing over a network and management blocks for
dynamically creating component blocks and event connections.

\subsection{Event Function Blocks} 
Event function blocks are some special basic function blocks
defined in the standard to ease the management of events.
These blocks are used for control, generation and detection
of events and can be used by other composite blocks. These
blocks are defined using an execution control chart or a
time sequence diagram. These blocks can be used for the
following purposes:
\begin{enumerate}
\item To split events to produce new events.
\item To allow events to be merged.
\item To permit the propagation of certain events.
\item To select between two or more events.
\item To delay an event by a given period.
\item To generate a stream of events at a regular interval.
\end{enumerate}

\section{Execution Model}
The execution of algorithms for basic function blocks is
invoked by the execution control portion of a function block
instance in response to events at event inputs. This
invocation takes the form of a request to the underlying
scheduling function of the associated resource to schedule
the execution of algorithm's operations. The model assumes
that the resource in which a function block exists provides
a scheduling function that ensures each phase of function
block execution occurs in the correct order and with the
correct priority.

The execution model specifies eight discrete timing points
as shown in Figure \ref{f:Execution_Model} that must occur
sequentially for the basic function block to operate. These
are:

\begin{description}
\item[$t_1$]: Values coming from external function blocks as
  input data are stable at the current block.
\item[$t_2$]: The event associated with the input values
  occurs.
\item[$t_3$]: The execution control function notifies the
  resource scheduling function that it has the input values
  and is ready to execute the algorithm.
\item[$t_4$]: The scheduling function begins the algorithm
  execution.
\item[$t_5$]: The algorithm processes input values, (and in
  some cases, internal variables) to create new output
  values that are written to the block's outputs.
\item[$t_6$]: The algorithm notifies the scheduling function
  that it has completed its execution.
\item[$t_7$]: The scheduling function invokes the execution
  control to generate the output events.
\item[$t_8$]: The execution control generates the
  appropriate output events at its event output ports. The
  event may be used by downstream blocks to signal that they
  may use the output values generated by this block.
\end{description}

\begin{figure}
  \begin{center}
    \epsfig{file=images/functionblocks/Execution_Control_Chart.eps, scale=.8}
    \caption[Execution Model]
            {Execution Model{\protect ~\cite{iec:614991:2000}}}
    \label{f:Execution_Model}
  \end{center}
\end{figure}

To ensure consistency, the algorithms must use the snapshot
of input variables that exists when the input event arrived.
Further the scheduling function must be able to detect the
arrival of input events at a faster rate than the rate at
which they can be handled by the block.