### 2004-12-06 doc/Simulation_and_Control_using_Supremica/introduction.tex
*** empty log message ***

% -*- TeX-master:"main" -*-

\chapter{Introduction}
\label{introduction}

%Define the problem.
Ever increasing competition forces companies to diversify
their product lines. To achieve customizable products the
companies require constantly changing and evolving
production plants. Such plants are called flexible
production plants. The plant is constantly adopted to the
products with goal of producing the highest quality products
in as little time as posibble. To achieve this the software
for the automatic control of such a plant has to be changed
along with the plant. The problem is that using today
established development methods this is impossible without
stoping of the production.

%Define other approaches, state of the art.

%Introduce Function Blocks

%Introduce Automation Objects